Black-swan events in animal populations

### 2016-11-17 text/ms_pnas.tex
Copyediting
Clarify the section on 0.5 threshold

For the Gompertz model,
we chose weakly informative priors for
all parameters (Fig.~S\figpriors, Supporting Information).
For the degrees of freedom parameter $\nu$,
we chose an exponential prior
with a rate parameter
of \(0.01\) truncated at values above two---a slightly less informative prior
than suggested by Ref.~\cite{fernandez1998}. We truncated the
distribution at \(2\) since
below $2$ there can be issues with computational stability.
This prior gives only a \(7.7\)\% probability that \(\nu < 10\)
but constrains the sampling sufficiently to avoid \(\nu\) drifting towards infinity. In
any case, for \(\nu > 20\) the $t$ distribution is almost indistinguishable from
the normal distribution (Fig.~1). Based on the shape of the $t$ distribution, we
chose the probability that \(\nu < 10\), Pr(\(\nu < 10\)), to define the
probability of heavy-tailed (black-swan) dynamics. When categorizing a population as heavy
tailed, we applied a 0.5 probability threshold. In the scenario where the data
are uninformative about heavy tails (e.g.~Fig.~1E, H), the posterior for
\(\nu\) will approximately match the prior (prior median \(= 71\), mean \(=
102\)) and a metric of Pr\((\nu < 10) > 0.5\) will not flag the
population as heavy tailed.

@@

\subsection{Alternative models and simulation testing}
We fit alternative population models to test if four key phenomena
systematically changed our conclusions. Autocorrelation has been suggested as
a reason for increased observed variability of abundance time series through
time, which could create apparent heavy tails \citep{inchausti2002};
therefore, we fit a model that included serial correlation in the residuals.
Additionally, previous work has modelled abundance or growth rates without
accounting for density dependence \citep{halley2002,segura2013}; therefore, we
fit a simpler model in which we assumed density independence.
This model is equivalent to a random walk with drift
model on log abundance, and therefore does not assume
that the timeseries is stationary.
Third, observation error could bias parameter estimates \citep{knape2012} or mask our
ability to detect heavy tails \citep{ward2007}; therefore, we fit a model
where we allowed for a fixed quantity of observation error ($0.2$ standard
deviations on a log scale). Fourth, the Gompertz model assumes that
population growth rate declines linearly with log abundance. Therefore, we
also fit an alternative model, the Ricker-logistic model, which assumes that
population growth rate declines linearly with abundance (Supporting Information).
% Our models also assume that their parameters
% (e.g. parameters representing productivity and density dependence)

>>>>

For the Gompertz model,
we chose weakly informative priors for
all parameters (Fig.~S\figpriors, Supporting Information).
For the degrees of freedom parameter $\nu$,
we chose an exponential prior
with a rate parameter
of \(0.01\) truncated at values above two---a slightly less informative prior
than suggested by Ref.~\cite{fernandez1998}. We truncated the
distribution at \(2\) since
below $2$ there can be issues with computational stability.
This prior gives only a \(7.7\)\% probability that \(\nu < 10\)
but constrains the sampling sufficiently to avoid \(\nu\) drifting towards infinity. In
any case, for \(\nu > 20\) the $t$ distribution is almost indistinguishable from
the normal distribution (Fig.~1). Based on the shape of the $t$ distribution, we
chose the probability that \(\nu < 10\), Pr(\(\nu < 10\)), to define the
probability of heavy-tailed (black-swan) dynamics. When categorizing a population as heavy
tailed, we applied a 0.5 probability threshold.
The data must therefore alter the posterior so that
more than 50\% of the posterior probability has $\nu < 10$ for a
population to be declared heavy tailed. 
If the data
are uninformative about heavy tails (e.g.~Fig.~1E, H), 
and the population model is otherwise a reasonable representation of the dynamics, 
the posterior for
\(\nu\) will approximately match the prior (prior median \(= 71\), mean \(=
102\)) and a metric of Pr\((\nu < 10) > 0.5\) will not flag any
populations as heavy tailed.

@@

\subsection{Alternative models and simulation testing}
We fit alternative population models to test if four key phenomena
systematically changed our conclusions. Autocorrelation has been suggested as
a reason for increased observed variability of abundance time series through
time, which could create apparent heavy tails \citep{inchausti2002};
therefore, we fit a model that included serial correlation in the residuals.
Additionally, previous work has modelled abundance or growth rates without
accounting for density dependence \citep{halley2002,segura2013}; therefore, we
fit a simpler model in which we assumed density independence.
This model is equivalent to a random walk with drift
on log abundance, and therefore does not assume
that the timeseries is stationary.
Third, observation error could bias parameter estimates \citep{knape2012} or mask our
ability to detect heavy tails \citep{ward2007}; therefore, we fit a model
where we allowed for a fixed quantity of observation error ($0.2$ standard
deviations on a log scale). Fourth, the Gompertz model assumes that
population growth rate declines linearly with log abundance. Therefore, we
also fit an alternative model, the Ricker-logistic model, which assumes that
population growth rate declines linearly with abundance (Supporting Information).
% Our models also assume that their parameters
% (e.g. parameters representing productivity and density dependence)
